---
output: html_document
params:
  state: "CA"
  district: 27
  endyear: 2015
  span: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

install.packages("acs", repos = "http://cran.us.r-project.org")
library(acs)
```

```{r api_key, include=FALSE}
source('acs_api_key.R')
```

```{r define_functions}
# Let's set up some geo objects to correlate to our district, state, and the whole country
geo_district <- geo.make(state=params$state, congressional.district=params$district)
geo_state <- geo.make(state=params$state)
geo_national <- geo.make(us=1)

# function to pull data out of the ACS tables
get_data <- function(table_number) {
  district_data <- estimate(acs.fetch(table.number=table_number, 
                                 geography=geo_district,
                                 endyear=params$endyear, 
                                 span = params$span, 
                                 col.names="pretty"))
  state_data <- estimate(acs.fetch(table.number=table_number, 
                                      geography=geo_state,
                                      endyear=params$endyear, 
                                      span = params$span, 
                                      col.names="pretty"))
  national_data <- estimate(acs.fetch(table.number=table_number, 
                                      geography=geo_national,
                                      endyear=params$endyear, 
                                      span = params$span, 
                                      col.names="pretty"))
  total_data <- rbind(district_data, state_data, national_data)
  return(total_data)
}

# function to graph data (to maintain consistency). Wants a matrix with 3 rows for national, state, and district data.
graph<- function(data) {
  barplot(data, beside=TRUE, col=c("dark blue", "light blue", "gray"))
}

# wrapper for our graph function that graphs proportions by row (to show percentages)
graph_as_percentage <- function(data) {
  graph(prop.table(data, 1)*100)
}
```


# `r params$state` `r params$district` District Overview

This report is automatically compiled from the Census Burea [American Community Survey](https://www.census.gov/programs-surveys/acs/). It is intended to provide basic demographic data for an individual congressional district. This report was complied based on the `r params$span` year span data ending in `r params$endyear`. Throughout this report there are graphs to show information about the district and comparing that data to state and national data. In all charts the district data is represented in dard blue, the state data in light blue, and the national data in gray.

## Income
```{r income_data}
income_raw <- get_data(table_number="B19001")
# combining data groups into chosen segements
income_data = cbind(income_raw[,2]+income_raw[,3]+income_raw[,4]+income_raw[,5],
                  income_raw[,6]+income_raw[,7]+income_raw[,8]+income_raw[,9]+income_raw[,10],
                  income_raw[,11]+income_raw[,12],
                  income_raw[,13],
                  income_raw[,14],
                  income_raw[,15],
                  income_raw[,16],
                  income_raw[,17])
# adding column names based on data grouping above
colnames(income_data) <- c("<$25K", "$25-50K", "$50-75K", "$75-100K", "$100-125K",
                       "$125-150K", "$150-200K", ">$200K")

graph_as_percentage(income_data)
```


## Ethnicity
```{r ethnicity_data}
ethnicity_raw = get_data(table_number="B02001")
ethnicity_data = cbind(ethnicity_raw[,2], 
                       ethnicity_raw[,3], 
                       ethnicity_raw[,4], 
                       ethnicity_raw[,5], 
                       ethnicity_raw[,6], 
                       ethnicity_raw[,7], 
                       ethnicity_raw[,8])
colnames(ethnicity_data) <- c("White", "Black", "Nat. Amer", "Asian", "Pac Island", "Other", "2+ races")
graph_as_percentage(ethnicity_data)
```

## Age
```{r age_data}
age_raw = get_data(table_number="B01001")
# adding together men and women and combining age groups
age_data <- cbind(age_raw[,3]+age_raw[,4]+age_raw[,27]+age_raw[,28], #<10
                 age_raw[,5]+age_raw[,6]+age_raw[,7]+age_raw[,29]+age_raw[,30]+age_raw[,31], #10-19
                 age_raw[,8]+age_raw[,9]+age_raw[,10]+age_raw[,11]+age_raw[,32]+age_raw[,33]+age_raw[,34]+age_raw[,35], #20-29
                 age_raw[,12]+age_raw[,13]+age_raw[,36]+age_raw[,37], #30-39
                 age_raw[,14]+age_raw[,15]+age_raw[,38]+age_raw[,39], #40-49
                 age_raw[,16]+age_raw[,17]+age_raw[,40]+age_raw[,41], #50-59
                 age_raw[,18]+age_raw[,19]+age_raw[,20]+age_raw[,21]+age_raw[,22]+age_raw[,23]+age_raw[,24]+age_raw[,25]+
                   +age_raw[,42]+age_raw[,43]+age_raw[,44]+age_raw[,45]+age_raw[,46]+age_raw[,47]+age_raw[,48]+age_raw[,49] #60+
                 )
# adding column names based on data grouping above
colnames(age_data) <- c("Under 10", "10-19","20-29", "30-39", "40-49", "50-59", "60+")
graph_as_percentage(age_data)
```

## Health Insurance
```{r health_insurance_data}
insurance_raw <- get_data(table_number="B27010")
# have  to add up coverage by age groups to get totals
insurance_data <- cbind(insurance_raw[,4]+insurance_raw[,20]+insurance_raw[,36]+insurance_raw[,53], # Employer
                        insurance_raw[,5]+insurance_raw[,21]+insurance_raw[,37]+insurance_raw[,54],# Direct Purchase
                        insurance_raw[,6]+insurance_raw[,22]+insurance_raw[,38]+insurance_raw[,55]+
                          insurance_raw[,7]+insurance_raw[,23]+insurance_raw[,39], # Medicare/Medicaid
                        insurance_raw[,8]+insurance_raw[,24]+insurance_raw[,40]+insurance_raw[,56]+
                            insurance_raw[,9]+insurance_raw[,25]+insurance_raw[,41]+insurance_raw[,57], # VA/Military
                        insurance_raw[,10]+insurance_raw[,26]+insurance_raw[,42]+insurance_raw[,58]) # multiple kinds
              
# setting columns names for data above
colnames(insurance_data) <- c('Employer', 'Individual', 'Medicaxx', "VA/Military", "Multiple")
graph_as_percentage(insurance_data)
```

## Employment
Note: this evaluates all individuals over 16 years of age.
```{r employment_data}
employment_raw <- get_data(table_number="B23025")
employment_data <- cbind(employment_raw[,4], #employed,
                         employment_raw[,6], #armed forces
                         employment_raw[,5], #unemployed
                         employment_raw[,7]) # not in labor force
colnames(employment_data) <- c("Employed", "Military", "Unemployed", "Out of LF")
graph_as_percentage(employment_data)
```
